---
title: "Agents for Scientists"
subtitle: "Pre-course work"
editor: source
execute:
  echo: false
  warning: false
editor_options:
  chunk_output_type: console
---

Thank you for joining this workshop. Your course instructor [{{< var author >}}]({{< var author-url >}}) is looking forward to meeting you.

We will meet on {{< var session.datetime >}} ({{< var session.timezone-label >}}) at {{< var session.event-location >}}.

This is the second of the two workshops in the series. The first, [{{< var prerequisite.title >}}]({{< var prerequisite.site >}}), ran on {{< var prerequisite.date >}} and is the foundation this day builds on.

## {{< fa circle-info >}} Background

In the first workshop you learned to use Git and GitHub so that every change to a project is versioned, documented, and open to review. This workshop puts a coding agent inside that same workflow.

The day runs the real work of publishing a paper (analysing data, making interpretations, writing documentation) under three conditions: as you do it now without agents, with an agent given minimal input, and with an agent given a proper plan. Comparing the three shows what the agent actually does for you, what it does badly, and what is left for you to do.

Because the work runs through Git and GitHub, every action an agent takes and every decision you make is recorded as you go. That record is what makes agent use traceable and declarable, which is the second half of the day: research integrity bodies are writing the disclosure rules right now, and this group will send them a response written from what we do in the room.

## {{< fa graduation-cap >}} Prior knowledge

Everyone taking part has been through the {{< var prerequisite.title >}} workshop, so this builds directly on it. You should be able to:

- clone a repository and open it as a project
- run the pull, stage, commit, push cycle
- create a branch and open a pull request
- review a colleague's pull request and merge it
- open an issue, assign it, and close it

If any of these are shaky, work through the [{{< var prerequisite.title >}} material]({{< var prerequisite.site >}}) again before the day. You do not need any prior experience with AI coding agents.

## {{< fa suitcase >}} What do I need to bring?

- a laptop, connected to wifi
- a web browser
- a pen and paper
- [your repository, with your own data already committed]{.highlight-yellow} (Step 1 below)

## {{< fa circle-question >}} What do I need to prepare?

[There is pre-work for this workshop]{.highlight-yellow}. It matters more than usual here: the day is built around *your* data, so a participant who arrives without it cannot take part in the main exercise.

Two tasks to start with, both due by [**{{< var due.repo >}}**]{.highlight-yellow}:

- [Step 1](pre-work/01-create-repository.qmd): Create your repository and add your data
- [Step 2](pre-work/02-join-element.qmd): Join the ai-playground chat room

Further preparation, including a pre-course survey and the tool setup, follows closer to the workshop.

If you would rather read before you do, the four [background pages](background/index.qmd) cover the security, literacy, transparency, and disclosure themes the day works through. They are optional.

```{r}
library(dplyr)
library(knitr)
library(readr)
```

## {{< fa calendar >}} Schedule

::: {.callout-note}
## Tentative

The schedule below is a working draft and still under active development. The shape of the day is settled; the individual blocks and timings will change.
:::

```{r}
#| tbl-colwidths: [25,75]

read_csv(here::here("data/tbl-01-agentsforsci-ghe-course-schedule.csv")) |>
  filter(day == 1) |>
  select(Time = time, Block = title) |>
  kable()

```

## {{< fa bullseye >}} Learning objectives

::: {.callout-note}
## Tentative

These objectives are still under active development and will be refined before the workshop.
:::

By the end of this workshop, participants will be able to:

1. **Run an agent inside the Git workflow**: give a coding agent a task, review what it produced as a pull request, and accept or reject the change.

2. **Judge the difference a plan makes**: compare the output of an agent given minimal input against one given a reviewed plan, and say what changed and why.

3. **Judge the difference documentation makes**: compare what an agent does with a plain CSV against the same data published as a data package.

4. **Declare agent use**: point to the record your workflow produced and say what it shows about which parts of the work an agent touched.

## {{< fa sun >}} Thanks

Thank you for working through these steps. See you at {{< var session.event-location >}}.
